---
title: "Friday R Tips"
output: html_document
date: "2023-06-03"
---

Last updated: 2 Jun 2023

I created this list of little r tricks that I've found interesting. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Setting a theme for all your plots

Below are a few ways to make sure all of the plots you make in a session match a specific theme, without having to write this out for each one.

### You can modify the default ggplot theme for the entire session
```{r change_default_theme, warning=FALSE, message=FALSE}
#The default theme
library(ggplot2)
ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + geom_bar()

#Change this to desired ggplot2 theme
theme_set(theme_void())
ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + geom_bar()
```

### You can also modify this default theme for each session
```{r modify_theme, warning=FALSE, message=FALSE}
theme_update(panel.grid.minor = element_line(colour = "red"))
ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + geom_bar()
```

### Or you can create your own theme
```{r own_theme, warning=FALSE, message=FALSE}
library(ggplot2)
Ugly_theme <- function (base_size = 12, base_family = "") {
    theme_bw() %+replace% 
    theme(
      panel.background = element_rect(fill = "black"),
      panel.border = element_rect(color = "white", fill = NA),
      axis.line = element_line(color = "white"),
      axis.ticks = element_line(color = "red"),
      axis.text = element_text(color = "red")
      )
}

ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + geom_bar() + Ugly_theme()
```

### There are also a couple of packages which can help

A really easy alternative is using the [ggpubfigs package](https://github.com/JLSteenwyk/ggpubfigs#faq), which has a great set of default themes and colourblind friendly colour schemes
```{r ggpubfigs, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggpubfigs)

# Simply add the friendly_pal() and theme_big_simple() elements to each ggplot command
ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + geom_bar() + scale_fill_manual(values = friendly_pal("ito_seven")) + theme_big_simple()
```

Another which stores your default theme between sessions is [thematic](https://rstudio.github.io/thematic/).
```{r thematic, warning=FALSE, message=FALSE, results='hide'}
library(ggplot2)
library(thematic)
library(RColorBrewer)

# Set your desired theme (use thematic_on in an r script not thematic_rmd())
thematic_rmd(
  bg = "#222222", fg = "white", qualitative = RColorBrewer::brewer.pal(8, "Dark2")
)

# Then just run this when you start a new r session
thematic::thematic_rmd

ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + geom_bar()

```