---
title: "Little R Tips"
output: html_document
date: "2023-06-03"
---


I created this list of little r tricks that I've found interesting. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Setting a theme for all your plots

Below are a few ways to make sure all of the plots you make in a session match a specific theme, without having to write this out for each one.

<br>

### (a) You can modify the default ggplot theme for the entire session
```{r change_default_theme, warning=FALSE, message=FALSE}
#The default theme
library(ggplot2)
ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + 
  geom_bar()

#Change this to desired ggplot2 theme
theme_set(theme_void())
ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + geom_bar()
```

### (b) You can also modify this default theme for each session
```{r modify_theme, warning=FALSE, message=FALSE}
theme_update(panel.grid.minor = element_line(colour = "red"))
ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + geom_bar()
```

### (c) Or you can create your own theme
```{r own_theme, warning=FALSE, message=FALSE}
library(ggplot2)
Ugly_theme <- function (base_size = 12, base_family = "") {
    theme_bw() %+replace% 
    theme(
      panel.background = element_rect(fill = "black"),
      panel.border = element_rect(color = "white", fill = NA),
      axis.line = element_line(color = "white"),
      axis.ticks = element_line(color = "red"),
      axis.text = element_text(color = "red")
      )
}

ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + geom_bar() + Ugly_theme()
```

<br>

### (d) There are also a couple of packages which can help

A really easy alternative is using the [ggpubfigs package](https://github.com/JLSteenwyk/ggpubfigs#faq), which has a great set of default themes and colourblind friendly colour schemes
```{r ggpubfigs, warning=FALSE, message=FALSE, echo=2:7}
library(ggplot2)
library(ggpubfigs)

# Simply add the friendly_pal() and theme_big_simple() elements to each ggplot command
ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + 
  geom_bar() + 
  scale_fill_manual(values = friendly_pal("ito_seven")) + theme_big_simple()
```

Another which stores your default theme between sessions is [thematic](https://rstudio.github.io/thematic/).
```{r thematic, warning=FALSE, message=FALSE, results='hide', echo=5:11}
library(ggplot2)
library(thematic)
library(RColorBrewer)

# Set your desired theme (use thematic_on in an r script not thematic_rmd())
thematic_rmd(
  bg = "#222222", fg = "white", qualitative = RColorBrewer::brewer.pal(8, "Dark2")
)

# Then just run this when you start a new r session
thematic::thematic_rmd

ggplot(mtcars, aes(factor(carb), fill=factor(cyl))) + geom_bar()

thematic::thematic_off()
```

<br><br>

## 2. Highlight regions of interest in a plot

Combining the `geom_rect` and `geom_text` functions in `ggplot2` is a really great way to highlight regions of interest in a plot. For example, let's look at the birth rate in Scotland in the 20^th^ century ([source](https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/vital-events/births/births-time-series-data)).

The `xmin` and `xmax` in `geom_rect` allow us to set the regions the highlighted box will cover. The `x` and `y` in `geom_text` allow us to position the labels, with `vjust` adjusting this vertically and `hjust` adjusting it horizontally.
```{r scotland_data, echo=FALSE, message=FALSE}
df <- read.csv("/Users/Andy/Desktop/birth_rate.csv")

df$Births <- as.numeric(gsub(",", "", df$Births))
```

```{r scotland_geom_rect, message=FALSE, echo=3:16}
library(ggplot2)
library(ggpubfigs)

ggplot(df, aes(x = Year, y = Births)) +
  geom_line() +
  geom_rect(aes(xmin = 1914, xmax = 1918, ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.002) +
  geom_rect(aes(xmin = 1939, xmax = 1945, ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.002) +
  geom_text(aes(x = 1918, y = max(Births), label = "WWI"), color = "black", vjust = 30, hjust = -0.3, size = 3.5) +
  geom_text(aes(x = 1945, y = max(Births), label = "WWII"), color = "black", vjust = 30, hjust = -0.3, size = 3.5) +
  xlab("Year") +
  ylab("No. of babies born in Scotland") +
  theme_simple()
```

<br><br>

## 3. Better model diagnostic plots
<br>
When we use models to examine the relationship between data, we make assumptions. For instance, a linear regression assumes that the relationship between the two variables is linear. Diagnostic plots are a key way of determining if the assumptions in our chosen models are being violated.


For instance, a linear regression to determine whether there is a correlation between education and fertility using the `Swiss` dataset in r.
```{r Base_diagnostic_plots, message=FALSE, results='hide'}
model_1 <- lm(Fertility~Education, data=swiss)

par(mfrow = c(2, 2))
plot(model_1)
```

However, these can be a little tricky to interpret. Here, the `check_model` function in the `performance` library comes in handy. It gives you a reminder of what to look for in each plot and makes each look a bit tidier:
```{r Better_diagnostic_plots, message=FALSE, results='hide'}
library(performance)
library(see)

check_model(model_1)
```

The package also contains functions to further dig into the results of these diagnostic plots. For example, it looks like the model's residuals might not be normally distributed and that the model's results may be influenced by outliers. We can check this using `check_outliers()` and `check_normality()`:
```{r homog_diagnostic_plots, message=FALSE}
check_outliers(model_1)
check_normality(model_1)
```
<br>